---
title: "Project: University Entrance Exam Analysis"
author: "Dol_R"
date: "8/24/2020"
output: 
  html_document:
    toc: true
    toc_depth: 3
    code_folding: hide
  pdf_document:
    toc: true
    toc_depth: 3
---

## Introduction and Key Takeaways
In Turkey, every year millions of students take the university entrance exam. After the announcement of results, participants list their university and department choices and they are placed according to their ranks.

<br>This analysis focuses on universities and departments popularity over the years. Hopefully, it would help future participants in their decision making process.

**Key Takeaways:**

* We analyzed University Exam Entrance data
* We explored the popularity trends of the universities and departments over the years
* We compared top selected universities and departments


## Summary of the Data and Explanations
Using **University Exam** data from  [*Hacettepe University's Website*](http://web.ee.hacettepe.edu.tr/osym), we obtained university results of years 2016-2020. Each year is on a separate Excel spreadsheet. Since every year some departments are opened and some are shut down, datasets will have different number of rows. There are 9 variables and more than 10000 rows for each dataset.

**university**: Name of the university
<br>**city**: University's location 
<br>**department**: Name of the department
<br>**type**: Type of the exam
<br>**quota**: Maximum number of participants to be accepted, upper bound for accepted_number
<br>**accepted_number**: Accepted number of participants to the selected university's selected department
<br>**lowest_score**: Lowest score of the accepted participant's scores
<br>**highest_score**: Highest score of the accepted participant's scores
<br>**lowest_ranking**: Lowest ranking of the accepted participant's scores, last accepted person's ranking

**Objectives:**

* Cleaning and manipulation of datasets
* Exploration of the popularity trends of the universities and departments
* Comparison by visualization
* Finding top selected universities and departments

### Data Preprocessing
In order to ease the reading process, functions `mani97` and `mani86` are created. These functions remove the columns related to the percentage of change regarding the previous year. They also rename the columns and remove the last empty rows. The raw data for some years have an additional column at the end, therefore two functions were needed.
```{r manipulation}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE) # applies to all chunks
mani97 <- function(data){
  newdata <- data %>% 
  rename(
    number = c(1),
    difference = c(2),
    university = c(3),
    city = c(4),
    department = c(5),
    type = c(6),
    quota = c(7),
    difference_quota = c(8),
    accepted_number = c(9),
    lowest_score = c(10),
    highest_score = c(11),
    lowest_ranking = c(12),
    difference_ranking = c(13)
  )  %>% select(-number, -difference, -difference_quota, -difference_ranking) %>% slice_head(n=nrow(data)-9)
  return(newdata)
}
mani86 <- function(data){
  newdata <- data %>% 
  rename(
    number = c(1),
    difference = c(2),
    university = c(3),
    city = c(4),
    department = c(5),
    type = c(6),
    quota = c(7),
    difference_quota = c(8),
    accepted_number = c(9),
    lowest_score = c(10),
    highest_score = c(11),
    lowest_ranking = c(12)
  )  %>% select(-number, -difference, -difference_quota) %>% slice_head(n=nrow(data)-9)
  return(newdata)
}
```
After loading the required libraries, we download the .xlsx file from our project repository. After reading it, we remove the file. 
```{r setup, results="hide"}
library(tidyverse)
library(lubridate)
library(tinytex)
library(readxl)
library(tidyr)
library(httr)
library(reshape2)
url<-'https://github.com/pjournal/boun01g-dol-r/blob/gh-pages/uni_exam_project/uni_exam.xlsx?raw=true'
GET(url, write_disk(tf <- tempfile(fileext = ".xlsx")))
raw_df20 <- read_excel(tf, sheet="20", skip=21)
raw_df19 <- read_excel(tf, sheet="19", skip=21)
raw_df18 <- read_excel(tf, sheet="18", skip=21)
raw_df17 <- read_excel(tf, sheet="17", skip=21)
raw_df16 <- read_excel(tf, sheet="16", skip=21)
file.remove(tf)
```

### Loading the Libraries and Datasets
We have 5 datasets for the last 5 years. Each dataset has 9 columns and they all have different number of rows. For example there are `r mani97(raw_df20) %>% summarise(exam20=n())` rows in dataset for the year 2020 and `r mani97(raw_df19) %>% summarise(exam19=n())` rows in dataset for the year 2019. Some departments or even universities may have been opened or closed. 

```{r datasets}
data2020 <- mani97(raw_df20)
data2020 <- mani97(raw_df20)
data2019 <- mani97(raw_df19)
data2018 <- mani86(raw_df18)
data2017 <- mani97(raw_df17)
data2016 <- mani97(raw_df16)

data2020 %>% summarise(exam20=n()) %>% mutate(data2019 %>% summarise(exam19=n())) %>% mutate(data2018 %>% summarise(exam18=n())) %>% mutate(data2017 %>% summarise(exam17=n())) %>% mutate(data2016 %>% summarise(exam16=n())) 
```

As an example, let's observe dataset for the year 2018. First four variables are categorical and the rest is numerical. We could also see that there are `r data2018 %>% summarise(exam18=n())` rows for this year.

```{r ex-dataset}
data2018 %>% arrange(desc(highest_score)) %>% glimpse()
```

## Exploratory Data Analysis
When the exam procedure changes, the types of grading systems also change. Therefore, there is no point of comparing `type` variable for different years. But it can be an identifier in each year.

### Top Selected Universities and Departments
Some universities' departments may be selected by the participants with highest scores in different types of grading systems. After grouping by `university` and `department`, we can order according to highest score and look at the first 10 rows. These are the universities and departments selected by the top participants. Let's see how they changed over the years. 

To do this, we count unique values and use `full_join`. Finally, we replace `NULL` values with zero, which means for that specific year, that university or department was not selected by the top 10 participants. 

```{r fnc, results="hide"}
score <- function(data){
  data %>% group_by(university, department) %>% select(-type, -city) %>% arrange(desc(highest_score)) %>% head(10) %>% mutate(department=replace(department, str_detect(department,"Tıp"), "Tıp")) %>% mutate(department=str_replace(department, "Burslu", "")) %>% mutate(department=str_replace(department, "İngilizce", "")) %>% mutate(department=str_replace(department, " \\([^()]{0,}\\)", "")) %>% mutate(department=str_replace(department, " \\([^()]{0,}\\)", "")) %>% ungroup()
}
dept <- function(data){
  data %>% count(department) %>% arrange(desc(n))
}
uni <- function(data){
  data %>% count(university) %>% arrange(desc(n))
}

```


```{r uni-dep}
top_uni <- full_join(uni(score(data2016)), uni(score(data2017)), by="university", suffix=c(".2016", ".2017")) %>% full_join(., uni(score(data2018)), by="university", suffix=c(".2016", ".2018"))
top_uni <- full_join(top_uni, uni(score(data2019)), by="university", suffix=c(".2018", ".2019"))
top_uni <- full_join(top_uni, uni(score(data2020)), by="university", suffix=c(".2019", ".2020"))
names(top_uni)[names(top_uni)=='n.2016']<-'year2016'
names(top_uni)[names(top_uni)=='n.2017']<-'year2017'
names(top_uni)[names(top_uni)=='n.2018']<-'year2018'
names(top_uni)[names(top_uni)=='n.2019']<-'year2019'
names(top_uni)[names(top_uni)=='n']<-'year2020'

top_dept <- full_join(dept(score(data2016)), dept(score(data2017)), by="department", suffix=c(".2016", ".2017")) %>% full_join(., dept(score(data2018)), by="department", suffix=c(".2016", ".2018"))
top_dept <- full_join(top_dept, dept(score(data2019)), by="department", suffix=c(".2018", ".2019"))
top_dept <- full_join(top_dept, dept(score(data2020)), by="department", suffix=c(".2019", ".2020"))
names(top_dept)[names(top_dept)=='n.2016']<-'year2016'
names(top_dept)[names(top_dept)=='n.2017']<-'year2017'
names(top_dept)[names(top_dept)=='n.2018']<-'year2018'
names(top_dept)[names(top_dept)=='n.2019']<-'year2019'
names(top_dept)[names(top_dept)=='n']<-'year2020'
```

```{r uni-dep-ggplot}
top_uni <- top_uni %>% replace_na(list(year2016 = 0, year2017 = 0, year2018 = 0, year2019 = 0, year2020 = 0))
top_dept <- top_dept %>% replace_na(list(year2016 = 0, year2017 = 0, year2018 = 0, year2019 = 0, year2020 = 0))

top_uni.long <- melt(top_uni)
plot_uni <- ggplot(top_uni.long, aes(x=value, y=university, col=variable)) +
  geom_point() + stat_smooth() + ggtitle("Top Universities") + xlab("Number of Participants") + ylab("Universities") + theme(plot.caption=element_text(hjust = 0.5))
plot_uni

top_dept.long <- melt(top_dept)
plot_dept <- ggplot(top_dept.long, aes(x=value, y=department, col=variable)) +
  geom_point() + stat_smooth() + ggtitle("Top Departments") + xlab("Number of Participants") + ylab("Departments") + theme(plot.caption=element_text(hjust = 0.5))
plot_dept
```

### Top Cities
```{r}
total_data <- full_join(data2016,data2017,by=c('university','department','city','type'),suffix = c('.2016','.2017'))%>%
  full_join(. ,data2018,by=c('university','department','city','type'))%>%
  full_join(. ,data2019,by=c('university','department','city','type'),suffix = c('.2018','.2019'))%>%
  full_join(. ,data2020,by=c('university','department','city','type'))
names(total_data)[names(total_data)=='type']<-'type.2020'
names(total_data)[names(total_data)=='quota']<-'quota.2020'
names(total_data)[names(total_data)=='accepted_number']<-'accepted_number.2020'
names(total_data)[names(total_data)=='lowest_score']<-'lowest_score.2020'
names(total_data)[names(total_data)=='highest_score']<-'highest_score.2020'
names(total_data)[names(total_data)=='lowest_ranking']<-'lowest_ranking.2020'
```

University and city distribution of the first thousand students in the university exam

```{r}
bin_tr<- total_data%>%
  filter(lowest_ranking.2020<1000)

tab2<-table(bin_tr$city,bin_tr$type.2020)
tab2

tab<-table(bin_tr$university,bin_tr$type.2020)
tab
```
University and city distribution of the first five thousand students in the university exam
```{r}
besbin_tr<- total_data%>%
  filter(lowest_ranking.2020<5000)

tab3<-table(besbin_tr$city,besbin_tr$type.2020)
tab3

  
tab4<-table(besbin_tr$university,besbin_tr$type.2020)
tab4
```

```{r}
besbin_tr%>%
  ggplot(aes(x = lowest_ranking.2020, fill = city)) +
  geom_density(alpha = 0.3)
  
```



### Quota Informations
The cities, universities and departments with the highest quotas in recent years can be seen from the tables below.

```{r}

all_years_data<-bind_rows(mutate(data2016,year=2016),
mutate(data2017,year=2017),
mutate(data2018,year=2018),
mutate(data2019,year=2019),
mutate(data2020,year=2020))

  all_years_data%>%
  group_by(year,department)%>%
  summarise(department_quota=sum(quota))%>%
   arrange(desc(department_quota))%>%head(20) 
  
 all_years_data%>%
  group_by(year,city)%>%
  summarise(city_quota=sum(quota))%>%
  arrange(desc(city_quota))%>%head(20)
    

  all_years_data%>%
  group_by(year,university)%>%
  summarise(university_quota=sum(quota))%>%
  arrange(desc(university_quota))%>%head(20)
  
```

### More...
  
  

## Conclusion



## References
[*Hacettepe University's Website*](http://web.ee.hacettepe.edu.tr/osym/).


