---
title: "Airbnb data for New York City 2019"
author: "Group Dol_R"
output: 
  html_document:
    toc: true
    toc_depth: 3
date: "8/20/2020"
---

## Introduction
Airbnb is an online marketplace for accommodations and it has more than 150 million users worldwide. In this report, the listing activity and metrics in New York City for 2019 will be analyzed according to **New York City Airbnb Open Data** data from  [*Kaggle*](https://www.kaggle.com/dgomonov/new-york-city-airbnb-open-data). 

## Loading Libraries and Data
```{r setup, message=FALSE, warning=FALSE, echo=TRUE}

library(readr)
library(tidyverse)
library(lubridate)
library(tinytex)
data <- read.csv("https://raw.githubusercontent.com/pjournal/boun01g-dol-r/gh-pages/NYC_assignment/AB_NYC_2019.csv", header = TRUE, check.names=TRUE)
```

## Overview and Manipulation of Data

### Variables
* __id__ : Listing ID
* __name__ : Name of the Listing
* __host_id__ : Host ID
* __host_name__ : Host Name
* __neighbourhood_group__ : Location
* __neighbourhood__ : Area
* __latitude__ : Latitude Coordinates
* __longitude__ : Longitude Coordinates
* __room_type__ : Listing Space Type
* __price__ : price (in US dollars)
* __minimum_nights__ : Minimum Nights
* __number_of_reviews__ : Total Number of Reviews
* __last_review__ : Date of the Last Review
* __reviews_per_month__ : 
* __calculated_host_listings_count__ : 
* __availability_365__ : 

<br> Numerical Variables: id, host_id, latitude, longitude, price, minimum_nights, number_of_reviews, last_review, reviews_per_month, calculated_host_listings_count, availability_365.All the numerical values can be considered as discrete.
<br> Categorical Variables: name, host_name, neighbourhood_group, neighbourhood, room_type 
<br> Since host_id identifies the host, there is no need for host_name and the respective column can be removed.  
```{r summary-manipulation}
data$last_review <- ymd(data$last_review)
data %>% glimpse()
```

### NA values
```{r null, warning=FALSE}
data %>% select(everything()) %>% summarise_all(funs(sum(is.na(.))))
```

## Analysis
Average price per night is `r data %>% summarise(avg_price = round(mean(data$price),2))`.
There are 3 different types of rooms which are `r unique(data$room_type)` . 
Average prices per room type  are : 
```{r room-type-mean, message=FALSE} 
data %>% select(price, room_type) %>% group_by(room_type)%>% summarise(mean(price))
``` 

### Summary of Dataset
There are `r data %>% summarise(count=n())` data points in total.

```{r summary, message=FALSE}
data %>% summary()
```

### Pie Plots for Categorical Variables
Lets show the ratios of Room Types and Neigbours with pie charts.
```{r pie-categorical, message=FALSE}
df_neigbour <- data.frame(data$neighbourhood_group)
df_neigbour <- unlist(df_neigbour)
df_neigbour <- data.frame(df_neigbour)
pie_chart <- ggplot(df_neigbour) + geom_bar(mapping = aes(x = "", fill=df_neigbour ), width = 1) + coord_polar("y") + 
  theme_void() # remove background, grid, numeric labels
pie_chart <- pie_chart + labs(x="", fill="Neigbours", title = "Dispersion of Neigbour Groups")
pie_chart
```

```{r pie-categorical2, message=FALSE}
df_rt <- data.frame(data$room_type)
df_rt <- unlist(df_rt)
df_rt <- data.frame(df_rt)
pie_chart2 <- ggplot(df_rt) + geom_bar(mapping = aes(x = "", fill=df_rt ), width = 1) + coord_polar("y") + 
  theme_void() # remove background, grid, numeric labels
pie_chart2 <- pie_chart2 + labs(x="", fill="Room Types", title = "Room Types Ratio")
pie_chart2
```

### Plots for Numerical Variables
```{r numerical, message=FALSE}

```

### Hosts with the Most Listings in NYC
```{r top-hosts, message=FALSE}

```

### Top 10 Neigbourhood
```{r top-neig, message=FALSE}

```

### Heatmap
```{r heatmap, message=FALSE}
#use the picture, x=latitude and y=longitude
#make shiny app here, slidebar for price and the heatmap changes according to neigbourhood_group or room_type

```

## References
- Dataset: **New York City Airbnb Open Data** from  [*Kaggle*](https://www.kaggle.com/dgomonov/new-york-city-airbnb-open-data). 



<br>To see Group Dol_R's Progress Journal, visit [Progress Journal website](https://pjournal.github.io/boun01g-dol-r/).




